---
title: "NYPD Shooting Data (Historic)"
author: "Ryan W"
date: "7/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Import Libraries
``` {r import libraries, echo=FALSE}
library(tidyverse)
```
## Importing Data
* The url for the data is https://data.cityofnewyork.us/resource/833y-fsy8.csv. Here we will set the url to variable name "urls" and 

``` {r importing, echo=FALSE}
url <- "https://data.cityofnewyork.us/resource/833y-fsy8.csv"
shootings <- read_csv(url)
```

## Tidying Data
Here we:
*remove unnecessary columns
*Remove rows with "NA"
*Here we rename columns
     
``` {r tidying, echo=FALSE}
shootings <- subset(shootings, select = -c(precinct:statistical_murder_flag, x_coord_cd:geocoded_column, incident_key:occur_time))
shootings <- na.omit(shootings)
shootings <- shootings %>% 
     rename(neighborhood = "boro",
     perp_age = "perp_age_group",
     vic_age = "vic_age_group")
```

## Summary of our Data
``` {r shootings summary}
summary(shootings)
```

## Analysis

### Number of shootings by neighborhood
``` {r shootings by neighborhood table}
shootings_neighborhood <- shootings %>% 
     group_by(neighborhood) %>%
     summarize(number_of_shootings=n()) %>%
     arrange(desc(number_of_shootings))
shootings_neighborhood 
```

### Number of shootings by sex

``` {r shootings by sex table}
shootings_perp_sex <- shootings %>% 
     group_by(perp_sex) %>%
     summarize(number_of_shootings=n()) %>%
     arrange(desc(number_of_shootings))
shootings_perp_sex 
```

## Visualization

### Number of shootings by neighborhood

``` {r shootings_neighborhood plot}
ggplot(shootings_neighborhood, aes(x=neighborhood, y=number_of_shootings)) + geom_col()
```
### Number of Shootings by Sex
``` {r shootings_perp_sex plot}
ggplot(shootings_perp_sex, aes(x=perp_sex, y=number_of_shootings)) + geom_col()
```

### Bias identification
Since I analyzed shootings by the perpetrators sex, I should note that I generally consider males as having a higher capacity for violence than females. Since the result of my analysis is aligned with my bias, you should keep that in mind. 

I don't know much about New York City, and I have no bias for the various neighborhoods there, so I don't think I had any bias affect my analysis.